---
title: "Mini Literate Analysis"
author: "Hailie Kittner"
format: html
---

## Introduction

### About this data

This is a dataset from annual sampling in the arctic ocean including CTD and oxygen data and seawater nutrients information from [Arctic Data Center](https://arcticdata.io/catalog/view/doi:10.18739/A25T3FZ8X).

### Setup

```{r}
library(readr)
library(here) # here function sets the default wd to be wherever the R project is 
```

### Read data

```{r}
#| warning: false
bg_chem <- read_csv("../data/BGchem2008data.csv") # read_csv from readr, instead of read.csv from base
# or
bg_chem <- read_csv(here("data/BGchem2008data.csv")) # does the same thing, but assuming we're not in the docs folder. You can't tab complete file names, but it's much more reproducible in the long run
```

Exploring our data

```{r}
#| eval: false
colnames(bg_chem)

str(bg_chem)

summary(bg_chem)

unique(bg_chem$Date)
```

## Analysis

### Calculate summary statistics

```{r}
nitrate <- mean(bg_chem$NO3)

nitrite <- mean(bg_chem$NO2)

amm <- mean(bg_chem$NH4)

phos <- mean(bg_chem$P)

```

### Calculate mean Redfield ratio
```{r}
ratio <- (nitrate+nitrite+amm)/phos
```


### Plot ratio

```{r}
plot(bg_chem$P, bg_chem$NO2 + bg_chem$NO3 + bg_chem$NH4)
```


## Conclusion

The Redfield ratio of this data is `r round(ratio, digits = 2)`
